---
title: "mate choice dataset"
author: "Luke Reding_ys"
date: "05/24/2019"
output:
  html_document:
    toc: true
    toc_float: true 
    geometry: margin=1.6in
    theme: "flatly"
    fontsize: 11pt
---

```{r eval = FALSE}

library(tidyverse)

```


```{r}

data_path <-"~/data_comgar_2019/matechoice_FRI_2019/data"

file_list <- dir(data_path,
                        pattern = "*.csv",
                        recursive = TRUE)
head(file_list)

```

```{r}

glue::glue("There are {length(file_list)} csv files in {data_path}.")

```

## create functions to extract useful information

Note filenames are named differently compared with the sociality data.

Mc = mate choice
Pre/post = pre/post numerosity
fid = fish ID (this will be followed by a number)
p1, p2, p3 = part (p1=5min background, p2=first   half, p3= second half)
Cornish= this is the scorer last name 
LargeMale_right= large male position 

```{r}
# FUNCTIONS
extractExperiment <- function(filename) {
  str_split(basename(filename), "_")[[1]][1]
}

extractPrePost <- function(filename) {
  str_split(basename(filename), "_")[[1]][2]
}

extractFishID <- function(filename) {
  str_split(basename(filename), "_")[[1]][3]
}

extractHalf <- function(filename) {
  str_split(basename(filename), "_")[[1]][4]
}

extractScorerName <- function(filename) {
  str_split(basename(filename), "_")[[1]][5]
}

extractLargeMalePosition <- function(filename) {
  str_split(basename(filename), "_")[[1]][7]
}

extractFishName <- function(filename) {
  str_split(basename(filename), "_")[[1]][8]
}

extractTreatment <- function(filename) {
  str_split(basename(filename), "_")[[1]][9]
}

# transits
numberTransits <- function(dataframe) {
  nrow(dataframe) - 1
}

# timeSpent
getTimeSpentInEachRegion <- function(dataframe) {
  
  # get the time the first starts
  startTime <- 0
  
  # calculate the time the trial ends
  endTime <- startTime + 300
  
  dataframe %>% 
    # make all the values of 'time' greater than endTime equal to endTime
    mutate(time = replace(time, time > endTime, endTime)) %>% 
    #take the differences. Any scoring after the end of the trial will get a 0 for timeSpent
    mutate(timeSpent = lead(time, default = 0) - time) %>% 
    # get rid of any zones where timeSpent = 0
    filter(timeSpent > 0) %>% 
    group_by(code) %>%
    # use summarise to distill information (take a sum or average)
    summarise(totalTime = sum(timeSpent)) 
}

getProportionOfTimeInEachRegion <- function(dataframe) {
  dataframe %>%
    getTimeSpentInEachRegion %>%
    group_by(code) %>%
       # use summarise to distill information (take a sum or average)
    summarise(totalTime = sum(timeSpent)) %>% 
     # use mutate to add columns
    mutate(proportionTime = totalTime / sum(totalTime)) 
   
}

# INTERACTION Stuff
getLatencyToEnterInteraction <- function(dataframe) {
  # get the first value of `time`. This is when the trial started
  timeOfTrialStart <- dataframe %>% slice(1) %>% pull(time)
  timeOfFirstInteractionZone <- dataframe %>% 
    filter(code == "interaction") %>% 
    slice(1) %>%
    pull(time)
  
  # if the fish never entered that zone, code as 600  (*== '=' testing equality)
  if(length(timeOfFirstInteractionZone) == 0) {
    timeOfFirstInteractionZone <- 300 + timeOfTrialStart
  }
  
  # take the difference
  timeOfFirstInteractionZone - timeOfTrialStart
  
  # timeOfFirstInteractionZone <- timeOfTrialStart
  }

#if first interaction zone is first code

FirstCodeIsInteractionCorrection <- function(dataframe) {
  getLatencyToEnterInteraction %>% 
    if(value == 0) {
      timeOfFirstInteraction <- timeOfTrialStart
    }
}

getLatencyToEnterLeftScreen <- function(dataframe) {
  leftScreeenLatency <- dataframe %>%
    filter(code == "left screen") %>%
    slice(1) %>%
    pull(time)%>%
    .[1]
  
  if_else(is.null(leftScreeenLatency), 300, leftScreeenLatency)
}

getLatencyToEnterRightScreen <- function(dataframe) {
  rightScreeenLatency <- dataframe %>%
    filter(code == "right screen") %>%
    slice(1) %>%
    pull(time) %>%
    .[1]
  
  if_else(is.null(rightScreeenLatency), 300, rightScreeenLatency)
}

getLatencyToEnterPlant <- function(dataframe) {
  plantLatency <- dataframe %>%
    filter(code == "plant") %>%
    slice(1) %>%
    pull(time)%>%
    .[1]
  
  if_else(is.null(plantLatency), 300, plantLatency)
}

clean_codenames <- function(data){
  # read in the csv
  # create a column for filename
  # create new variables
  # get the transits and proportions  
  #'Mutate = adds or overwrites columns, 
  # map_dbl returns a floating decimal..
  # map replaces a for loop
data %>%
  mutate(code=case_when(
    code == "a" ~ "right screen",
    code == "s" ~ "right side", 
    code == "d" ~ "center",
    code == "m" ~ "plant",
    code == "o" ~ "left side",
    code == "p" ~ "left screen",
    TRUE ~ "unknown"
  ))
}

```

Let's make sure these have the expected behavior:


Previously we had a function that calculated the latency to enter the interaction zone. 
Since we don't have an interaction zone, I've deleted this function in this document.

```{r eval = FALSE}

full_dataframe <- tibble(filename = file_list) %>%
  mutate(file_contents = map(filename, ~ read_csv(file.path(data_path, .)))) %>%
  mutate(file_contents = map(file_contents, clean_codenames)) %>%
  mutate(number_transits = map_dbl(file_contents, numberTransits),
         timeSpent = map(file_contents, getTimeSpentInEachRegion),
         latency_plant = map_dbl(file_contents, getLatencyToEnterPlant),
         latency_right_screen = map_dbl(file_contents, getLatencyToEnterRightScreen),
         latency_left_screen = map_dbl(file_contents, getLatencyToEnterLeftScreen)

           
         # latencyPlant = map(file_contents, getLatencyToEnterPlant),
         # latencyRightScreen = map(file_contents, getLatencyToEnterRightScreen),
         # latencyLeftScreen = map(file_contents, getLatencyToEnterLeftScreen)
  ) %>%
  unnest(timeSpent) %>%
  # complete(filename, code, fill = list(totalTime = 0, proportionTime = 0)) %>% 
  # I'm getting rid of this line because it cause more harm than good. 
  # We will fill these values with 0 later on in the process
  rowwise() %>%
  mutate(experiment = extractExperiment(filename),
         fishID = extractFishID(filename),
         fishName = extractFishName(filename),
         treatment = extractTreatment(filename),
         half = extractHalf(filename),
         scorer = extractScorerName(filename),
         largeMalePosition = extractLargeMalePosition(filename),
         prePost = extractPrePost(filename)
         
  ) %>%
  ungroup # undos rowwise(), which we don't need anymore
sample_n(full_dataframe, 10)

View(full_dataframe)

```

Right now, there's a row for each zone in each of the fish's trials for each observer.

We might want the data in a different format, where there's a single row for each trial for each observer. 
When this is the case, look at the `tidyr` package.

```{r eval = FALSE}

alternate_df <- full_dataframe %>%
  select(filename, code, totalTime, number_transits, scorer, half, experiment,fishID, largeMalePosition, %>%
           prePost, latency_plant, latency_right_screen, latency_left_screen) %>%
  spread(code, totalTime) %>% # create columns for each zone
  replace(., is.na(.), 0) # replace NAs (where the fish didn't enter a zone) with 0

View(alternate_df)
sample_n(alternate_df, 6)

```

###REQUEST 21 May 2019.  We're not sure where this goes, we would like to take the Median 
across all the 3-5 scorers that have scored the same video.  So for instance, where experiment, 
fishID, side,half, camera were identical... and the only difference was ScorerName.  

```{r eval = FALSE}
medians_result <- full_dataframe %>%
  group_by(experiment, fishID, half, code, prePost, largeMalePosition) %>%
  summarise_at(c("totalTime", "number_transits"), median, na.rm = TRUE)

medians_result %>% head
```


```{r eval = FALSE}

medians_result <- full_dataframe %>%
  group_by(experiment, fishID, half, code, prePost, largeMalePosition) %>%
  summarise_at(c("totalTime", "number_transits"), median, na.rm = FALSE)

medians_result %>% head
```

###REQUEST 21 May 2019: AFTER we calculate the medians, we then want to AVERAGE 
between the 1st and 2nd halves (extractHalf) of the medians of proportion time in each zone. 
{We reasoned that because different students scored 1st and 2nd halves, it didn't makes sense 
to pool before we caluclated the median}

I'm a bit confused here


```{r eval = FALSE}

# stop here and inspect
medians_result %>%
  select(experiment, fishID, half, code, prePost, largeMalePosition, proportionTime) %>%
  spread(half, proportionTime) %>%
  filter(fishID == "fid1") %>%
  View
  
```

```{r eval = FALSE}

medians_result %>%
  select(experiment, fishID, half, code, prePost, largeMalePosition, proportionTime) %>%
  spread(half, proportionTime) %>%
  rowwise() %>%
  mutate(averaged = mean(c(p1, p2, p3), na.rm = TRUE)) %>%
  ungroup %>%
  replace_na(list(p1 = 0, p2 = 0, p3 = 0)) # fill NAs with 0

```



